---
title: "Results"
author: "Lara Wootton"
date: "26/07/2020"
output: pdf_document
---

The BEAST analysis



The terminal branches of the Lowlands clade showed the highest rate of estimated non-synonymous (dN) and synonymous (dS) changes, followed by those of the Ramosa and Setacea clades (Fig. ). Conversely, the terminal branches of Setacea clade had a significantly greater $\omega$ than the Lowlands clade (p <0.05, Fig), although the estimated $\omega$ for all clades was low at <0.5.  For each of the three clades, a branch model allowing the $\omega$ of foreground branches to differ from background $\omega$ fit the data better than a model with a single $\omega$ across the tree. Both the Ramosa and Setacea clades showed a higher $\omega$ than the background rate, while Lowlands clade had a lower $\omega$ compared to the rest of the tree. 










```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = T, echo=FALSE, warning=FALSE, message=FALSE)

require(kableExtra)
require(motmot)
require(ape)
require(wesanderson)
require(tidyverse)
require(here)
require(BAMMtools)

```

\renewcommand{\arraystretch}{1.5}
```{r date_table}

#Read in trees

tree1 <- read.nexus(here("Data", "run1_full_annotated.tre"))
tree2 <- read.nexus(here("Data", "run2_full_annotated.tre"))
tree3 <- read.nexus(here("Data", "run3_full_annotated.tre"))
tree4 <- read.nexus(here("Data", "run4_full_annotated.tre"))
tree5 <- read.nexus(here("Data", "run5_full_annotated.tre"))
tree6 <- read.nexus(here("Data", "run6_full_annotated.tre"))
tree7 <- read.nexus(here("Data", "run7_full_annotated.tre"))
tree8 <- read.nexus(here("Data", "run8_full_annotated.tre"))
tree9 <- read.nexus(here("Data", "run9_full_annotated.tre"))
tree10 <- read.nexus(here("Data", "run10_full_annotated.tre"))
tree11 <- read.nexus(here("Data", "Redo_good_full_b20.tre"))
tree12 <- read.nexus(here("Data","sorta_bad_full_annotated.tre"))

all_trees <- c(tree1, tree2,tree3, tree4,tree5, tree6, tree7, tree8, tree9, tree10, tree11, tree12)

tree_ages <- matrix(nrow = 12, ncol = 7)

for (i in 1:12) {
    
    tree <- ladderize(all_trees[[i]], right = T)
    tree <- TreeTools::Renumber(tree)
    
    age <- dispRity::tree.age(tree)
    tree_ages[i, 2] <- round(as.numeric(age[c(43),1]), 3)
    tree_ages[i, 3] <- round(as.numeric(age[c(73),1]), 3)
    tree_ages[i, 4] <- round(as.numeric(age[c(47),1]), 3)
    tree_ages[i, 5] <- round(as.numeric(age[c(48),1]), 3)
    tree_ages[i, 6] <- round(as.numeric(age[c(66),1]), 3)
    tree_ages[i, 7] <- round(as.numeric(age[c(74),1]), 3)
    tree_ages[i, 1] <- i
}

tree_ages[11, 1] <- "Sortadate 'best'"
tree_ages[12, 1] <- "Sortadate 'worst'"
tree_ages <- rbind(tree_ages, c("Mean age", 
                    round(mean(as.numeric(tree_ages[,2]), na.rm = T), 3), 
                    round(mean(as.numeric(tree_ages[,3]), na.rm = T), 3), 
                    round(mean(as.numeric(tree_ages[,4]), na.rm = T), 3),
                    round(mean(as.numeric(tree_ages[,5]), na.rm = T), 3),
                    round(mean(as.numeric(tree_ages[,6]), na.rm = T), 3),
                    round(mean(as.numeric(tree_ages[,7]), na.rm = T), 3)
                    ))


kable(tree_ages,
      caption = "Median node ages (Ma) for the crown node of the Setacea, Ramosa, Succulent Karoo (SK) and Lowlands clades for BEAST trees either each constructed from 25 randomly chosen supercontig loci from the ninety coverage dataset, or from the gene 'shopping' method of Sortadate (Smith et al. 2018).",
      format = "latex",
      align = "r",
      linesep = "",
      digits = c(1,3,3,3,3,3,3),
      booktabs = T,
      col.names = c("Tree", "Root", "Microlaena", "Lowlands","SK", "Ramosa", "Setacea"))  %>%
  row_spec(12, hline_after = T) %>%  
  kable_styling(latex_options = c("striped", "hold_position"),
                    full_width = F,
                  position = "left") 

```

```{r bamm_tree, results = 'hide'}
tree5 <- read.tree(here("Data", "tree5_BAMM.tre"))
x <- getEventData(tree5, here("Data", "tree5_event_data.txt"), burnin = 0.1)

plot.bammdata(x, border = F, 
                      pal = "BuDkOr", 
                      lwd = 2.5, 
                      cex=0.5,
                      labels = T,
                      legend = T)

#css <- credibleShiftSet(x,threshold=5, expectedNumberOfShifts= 1)
#plot.credibleshiftset(css, border = F, 
#                      pal = "BuDkOr", 
##                      lwd = 2.5, 
#                      cex=0.7,
#                      labels = T,
#                      legend = T)
```


```{r bamm_clade_rates, results='hide'}

rup <- vector(length = 10)
ram <- vector(length = 10)
low <- vector(length = 10)
sk <- vector(length = 10)

for (i in 1:10) {
    tree <- read.tree(here("Data", paste0("tree", i, "_BAMM.tre" )))
    x <- getEventData(tree, here("Data", paste0("tree", i, "_event_data.txt" )), burnin = 0.1)
    
    rup_rates <- getCladeRates(x, node=70)
    rup[i] <- mean(rup_rates$lambda)
    ram_rates <- getCladeRates(x, node=63)
    ram[i] <- mean(ram_rates$lambda)
    low_rates <- getCladeRates(x, node=44)
    low[i] <- mean(low_rates$lambda)
    sk_rates <- getCladeRates(x, node=45)
    sk[i] <- mean(sk_rates$lambda)
    
    
}

clades <- data.frame(Rate =c( unlist(low),
                      unlist(sk),
                      unlist(ram),
                      unlist(rup)),
                      Clade = c(rep("Lowlands", 10),
                                rep("S. Karoo", 10),
                                rep("Ramosa", 10),
                                rep("Setacea", 10)))
                     


my_theme <- theme_bw() +
    theme(panel.grid.minor = element_blank(),
          panel.grid = element_blank(),
          legend.key = element_rect(fill = NA, color = "grey"),
          legend.background = element_blank(),
          axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          axis.ticks.x = element_blank(),
          legend.text = element_text(size = 15)
    )

ggplot(clades, aes(Clade, Rate, fill = Clade)) +
    geom_boxplot() +
    my_theme +
        scale_fill_manual(values=wes_palette(n=4, name="Moonrise2"), 
                          labels = c("Lowlands", "Ramosa", "S. Karoo",                                             "Setacea"),
                          name = " ")

```


```{r, cache=T}

trait_data <- read.csv(here("Data", "trait_dataV4.csv"))

trait_dat <- trait_data %>% 
    mutate(inflor_position = Plant_height/Veg_height,
           relative_sheath_length = Sheath_length/Internode_dist,
           relative_leaf_length = Leaf_length/Sheath_length,
           glume_enclosure = Glume_length_outer/Spikelet_length,
           glume_ratio = Glume_length_outer/Glume_length_inner,
           spikelet_shape = Spikelet_length/Spikelet_width,
           lemma_ratio = Lemma_outer_length/Lemma_inner_length)



#make labels the same as tips for pop level samples
pop_dat_id <- paste(trait_dat$Collection[1:315], trait_dat$Sample[1:315], sep = "-") 
pop_dat_id[1] <- "LW3-1"
pop_dat_id[315] <- NA #Don't want to use this microlaena
pop_dat_id[144] <- NA #duplicated id

low_dat_id <- as.character(trait_dat$Collection[316:339]) #pull out other lowland names
low_dat_id[16] <- "1628-3" #Need to use 695 data for microleana, so rename for ease of use
low_dat_id[2] <- "1625-u" 
low_dat_id[6] <- "1554" #typo


id <- c(pop_dat_id, low_dat_id)

trait_dat$ID <- id

#Filter individuals that are in the tree

tips <- tree1$tip.label
date_trait <- trait_dat %>% filter(ID %in% tips)


#Add column coded for elevation

date_trait <- date_trait %>% 
    mutate(Elevation = case_when(
        ID %in% c("LW3-1", "1649-1","1592-2", 
                  "1635-2", "1576-1", "1622-1", "1611-2") ~ 1,
        ID %in% c("1605-3", "1572-2", "1594-2", "1600-2", "1597-2","1655-1","1657-2",
                  "1644-2","1621-4", "1602-3", "1629-3") ~ 2,
        ID %in% c("1625-u", "1555", "1556", "1554", "1519", "1516", "1552", "1525", "1543", "1542", "1532", "1550", "1549", "1551", "1547", "1534",
                  "1526", "1535", "1544", "1523") ~ 0,
        ID %in% c("1628-3", "1646") ~ 3
    ))

row.names(date_trait) <- date_trait$ID


#Loop

rate_data <- matrix(ncol = 5, nrow = 1000)
n <- 0


for (i in c(4:8, 11:21)) {
    for (j in 1:length(all_trees)) {
        
        n <- n + 1
        
        tree <- ladderize(all_trees[[j]], right = T)
        
        tree <- TreeTools::Renumber(tree)
        
        sortedData <- sortTraitData(phy = tree, y = date_trait, 
                                    data.name = names(date_trait[i]),
                                    log.trait = T)
        
        phy <- sortedData$phy
        trait <- sortedData$trait
        
        model_h <- transformPhylo.ML(phy = phy, y = trait, 
                                   model = "clade", nodeIDs = c(74))
        model_m <- transformPhylo.ML(phy = phy, y = trait, 
                                     model = "clade", nodeIDs = c(66))
        model_l <- transformPhylo.ML(phy = phy, y = trait, 
                                     model = "clade", nodeIDs = c(47))
        
        rate_data[n,1] <- j
        rate_data[n,2] <- model_l$Rates[2]
        rate_data[n,3] <- model_m$Rates[2]
        rate_data[n,4] <- model_h$Rates[2]
        rate_data[n,5] <- names(date_trait[i])
    }
    
}

rate_data <- as.data.frame(rate_data)
names(rate_data) <- c("Tree", "Rate_low", "Rate_mid", "Rate_high", "Trait")
rate_data <- na.omit(rate_data)
rate_data[,4] <- as.numeric(as.character(rate_data[,4]))
rate_data[,3] <- as.numeric(as.character(rate_data[,3]))
rate_data[,2] <- as.numeric(as.character(rate_data[,2]))
rate_data[,1] <- as.numeric(as.character(rate_data[,1]))

rate_data <- rate_data %>% mutate(Var = case_when(
    Trait == "Veg_height" ~ "Vegetative height" ,
    Trait == "Spikelet_width" ~ "Spikelet width",
    Trait == "Spikelet_no" ~ "Spikelet number",
    Trait == "Spikelet_length"~ "Spikelet length",
    Trait == "Sheath_length" ~ "Sheath length",
    Trait == "Plant_height" ~ "Plant height",
    Trait == "Pedicel_length_epu" ~ "Pedicel length",
    Trait == "Lemma_outer_length" ~ "Outer lemma length",
    Trait == "Lemma_inner_length" ~ "Inner lemma length",
    Trait == "Leaf_width" ~ "Leaf width",
    Trait == "Leaf_length" ~ "Leaf length",
    Trait == "Internode_dist" ~ "Internode distance",
    Trait == "Inflor_length" ~ "Inflorescence length",
    Trait == "Glume_length_outer" ~ "Outer glume length",
    Trait == "Glume_length_inner" ~ "Inner glume length",
    Trait == "Culm_diam" ~ "Culm diameter"
    
))

rate_data$Var <- factor(rate_data$Var, levels = c(
    "Plant height",
    "Vegetative height",
    "Leaf length",
    "Leaf width",
    "Sheath length",
    "Internode distance",
    "Culm diameter",
    "Inflorescence length",
    "Spikelet number",
    "Pedicel length",
    "Spikelet length",
    "Spikelet width",
    "Outer glume length",
    "Inner glume length",
    "Outer lemma length",
    "Inner lemma length"
))

slow_rates_high <- rate_data %>% filter(Rate_high < 1)
slow_rates_mid <- rate_data %>% filter(Rate_mid < 1)
slow_rates_low <- rate_data %>% filter(Rate_low < 1)
```

```{r trait_rate, fig.width=8, cache=F}
library(viridis)
library(ggnewscale)

a <- ggplot(rate_data, aes(y = Var, x = Tree, fill = Rate_high)) +
    geom_tile() +
    scale_fill_viridis(option = "magma", direction = -1, limits = c(1,10)) +
    ylab("Setacea clade") +
    new_scale_fill() + 
    geom_tile(data = slow_rates_high, aes(y = Var, x = Tree, fill = Rate_high))+
                  scale_fill_distiller(palette = "YlGnBu", limits = c(0, 1), breaks = c(0,0.3,0.6,0.9)) +
    theme(axis.title = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          panel.background = element_blank(),
          plot.title = element_text(hjust = 0.5)) +
    scale_x_continuous(breaks = 1:12) +
    labs(title = "Setacea clade")

b <- ggplot(rate_data, aes(y = Var, x = Tree, fill = Rate_mid)) +
    geom_tile()+
    scale_fill_viridis(option = "magma", direction = -1, limits = c(1,10))+
    ylab("Ramosa clade")+
    new_scale_fill() +
    geom_tile(data = slow_rates_mid, aes(y = Var, x = Tree, fill = Rate_mid)) +
    scale_fill_distiller(palette = "YlGnBu", limits = c(0, 1), breaks = c(0,0.3,0.6,0.9)) +
    theme(axis.title.y = element_blank(),
          axis.title.x = element_text(colour = "black", size = 13),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          panel.background = element_blank(),
          plot.title = element_text(hjust = 0.5)) +
    scale_x_continuous(breaks = 1:12) +
    labs(title = "Ramosa clade")

c <- ggplot(rate_data, aes(y = Var, x = Tree, fill = Rate_low)) +
    geom_tile()+
    scale_fill_viridis(option = "magma", direction = -1, limits = c(1,10), breaks = c(1,5,10))+
    ylab("Lowlands clade")+
    labs(fill = "ML rate") +
    new_scale_fill() +
    geom_tile(data = slow_rates_low, aes(y = Var, x = Tree, fill = Rate_low)) +
    scale_fill_distiller(palette = "YlGnBu",limits = c(0, 1), breaks = c(0,0.3,0.6,0.9)) +
    theme(axis.ticks.y = element_blank(),
          axis.title = element_blank(),
          axis.text.y = element_text(size = 12, colour = "black"),
          #axis.text.x = element_blank(),
          panel.background = element_blank(),
          plot.title = element_text(hjust = 0.5)) +
    scale_x_continuous(breaks = 1:12) +
    labs(title = "Lowlands clade", fill = " ")

ggpubr::ggarrange(c,b,a, 
                  nrow = 1, 
                  common.legend = T,
                  legend = "right",
                  widths = c(1,0.6,0.6),
                  align = "h")

```

```{r contmap, results = 'hide', fig.width=8, fig.height=8,cache=F}

require(phytools)
par(mfrow = c(2,2))

tree <- drop.tip(tree1, c("M1_S1_L001", "Oryza_sativa"))

#veg height
vegHeight <- date_trait$Veg_height
names(vegHeight) <- date_trait$ID
a <- contMap(tree, log(vegHeight), plot = F)

plot(a, legend=FALSE,
     fsize = 0.5,
     ylim=c(1-0.09*(Ntip(a$tree)-1),Ntip(a$tree)),
    mar=c(5.1,0.4,0.4,0.4))
add.color.bar(15, a$cols,title="log Veg. height", 
    lims=a$lims,digits=2,prompt=FALSE,x=0,
    y=1-0.08*(Ntip(a$tree)-1),lwd=4,fsize=1,subtitle="")

#internode dist
inter <- date_trait$Internode_dist
names(inter) <- date_trait$ID
b <- contMap(tree, log(inter), plot = F)

plot(b, legend=FALSE,
     fsize = 0.5,
     ylim=c(1-0.09*(Ntip(b$tree)-1),Ntip(b$tree)),
    mar=c(5.1,0.4,0.4,0.4))
add.color.bar(15, b$cols,title="log Internode dist.", 
    lims=b$lims,digits=2,prompt=FALSE,x=0,
   y=1-0.08*(Ntip(a$tree)-1),lwd=4,fsize=1,subtitle="")

#Glume inner a
gl_in <- date_trait$Glume_length_inner
names(gl_in) <- date_trait$ID
c <- contMap(tree, log(gl_in), plot = F)

plot(c, legend=FALSE,
     fsize = 0.5,
     ylim=c(1-0.09*(Ntip(c$tree)-1),Ntip(c$tree)),
    mar=c(5.1,0.4,0.4,0.4))
add.color.bar(15, c$cols,title="log Inner glume len.", 
    lims=c$lims,digits=2,prompt=FALSE,x=0,
   y=1-0.08*(Ntip(c$tree)-1),lwd=4,fsize=1,subtitle="")

#Lemma inner
lem_in <- date_trait$Lemma_inner_length
names(lem_in) <- date_trait$ID
d <- contMap(tree, log(lem_in), plot = F)

plot(d, legend=FALSE,
     fsize = 0.5,
     ylim=c(1-0.09*(Ntip(d$tree)-1),Ntip(d$tree)),
    mar=c(5.1,0.4,0.4,0.4))
add.color.bar(15, d$cols,title="log Inner lemma len.", 
    lims=c$lims,digits=2,prompt=FALSE,x=0,
   y=1-0.08*(Ntip(d$tree)-1),lwd=4,fsize=1,subtitle="")




```


```{r disparity, cache=T, fig.width = 8, fig.height=8, results= 'hide', fig.cap = "Trait disparity within clades. Letters signify significance based on Leven's test for equal variance"}

my_theme <- theme_bw() +
    theme(panel.grid.minor = element_blank(),
          panel.grid = element_blank(),
          legend.key = element_rect(fill = NA, color = "grey"),
          legend.background = element_blank(),
          axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          axis.ticks.x = element_blank(),
          legend.text = element_text(size = 15)
    )


univariate_boxplot_trait_data <- function(x) {
    ggplot(func_traits, aes_string("Elevation", x)) +
        geom_boxplot(aes(fill = Elevation),
                     width = 0.6) +
        my_theme +
        scale_fill_manual(values=wes_palette(n=3, name="Moonrise2"), 
                          labels = c("Lowlands", "Ramosa", "Setacea"),
                          name = " ")
        
}

levens <- function(x) {
    
    func_traits2 <- func_traits %>% select(x, Elevation)
    
    df <- func_traits2 %>% group_by(Elevation) %>% 
       mutate(dat.med = median(!! sym(x))) %>% 
       mutate(dat_med_res = abs(!! sym(x) - dat.med))
    
    aov(df$dat_med_res~df$Elevation)
    
}


library(multcompView)
# I need to group the treatments that are not different each other together.
generate_label_df <- function(TUKEY, variable){
    
    # Extract labels and factor levels from Tukey post-hoc 
    Tukey.levels <- TUKEY[[variable]][,4]
    Tukey.labels <- data.frame(multcompLetters(Tukey.levels)['Letters'])
    
    #I need to put the labels in the same order as in the boxplot :
    Tukey.labels$treatment=rownames(Tukey.labels)
    Tukey.labels=Tukey.labels[order(Tukey.labels$treatment) , ]
    return(Tukey.labels)
}

func_traits <- date_trait[,c(4:8, 11:21, 33)]
#func_traits$Elevation <- as.numeric(levels(func_traits$Elevation))[func_traits$Elevation]
func_traits <- func_traits %>% filter(Elevation < 3)
func_traits$Elevation <- as.factor(func_traits$Elevation)


#Plant height
try <- levens("Plant_height")
summary(try)
hsd <- TukeyHSD(try)

LABELS <- generate_label_df(hsd, 'df$Elevation')

output <- data.frame(LABELS)
a <- univariate_boxplot_trait_data("Plant_height") +
    ylab("Plant height (mm)") +
    geom_text(data = output, 
              aes(x = treatment, y = rep(1.1*max(func_traits$Plant_height), 3), label = Letters),
              size = 6)

    
#Veg height
try <- levens("Veg_height")
summary(try)
hsd <- TukeyHSD(try)

LABELS <- generate_label_df(hsd, 'df$Elevation')

output <- data.frame(LABELS)
b <- univariate_boxplot_trait_data("Veg_height") +
    ylab("Vegetative height (mm)") +
    geom_text(data = output, 
              aes(x = treatment, y = rep(1.1*max(func_traits$Veg_height), 3), label = Letters),
              size = 6)

#Leaf width
try <- levens("Leaf_width")
summary(try)
hsd <- TukeyHSD(try)

LABELS <- generate_label_df(hsd, 'df$Elevation')

output <- data.frame(LABELS)
c <- univariate_boxplot_trait_data("Leaf_width") +
    ylab("Leaf width (mm)") +
    geom_text(data = output, 
              aes(x = treatment, y = rep(1.1*max(func_traits$Leaf_width), 3), label = Letters),
              size = 6)

#Leaf length
try <- levens("Leaf_length")
summary(try)
hsd <- TukeyHSD(try)

LABELS <- generate_label_df(hsd, 'df$Elevation')

output <- data.frame(LABELS)
d <- univariate_boxplot_trait_data("Leaf_length") +
    ylab("Leaf length (mm)") +
    geom_text(data = output, 
              aes(x = treatment, y = rep(1.1*max(func_traits$Leaf_length), 3), label = Letters),
              size = 6)

#Sheath
try <- levens("Sheath_length")
summary(try)
hsd <- TukeyHSD(try)

LABELS <- generate_label_df(hsd, 'df$Elevation')

output <- data.frame(LABELS)
e <- univariate_boxplot_trait_data("Sheath_length") +
    ylab("Sheath length (mm)") +
    geom_text(data = output, 
              aes(x = treatment, y = rep(1.1*max(func_traits$Sheath_length), 3), label = Letters),
              size = 6)

#internode

try <- levens("Internode_dist")
summary(try)
hsd <- TukeyHSD(try)

LABELS <- generate_label_df(hsd, 'df$Elevation')

output <- data.frame(LABELS)
f <- univariate_boxplot_trait_data("Internode_dist") +
    ylab("Internode distance (mm)") +
    geom_text(data = output, 
              aes(x = treatment, y = rep(1.1*max(func_traits$Internode_dist), 3), label = Letters),
              size = 6)

#culm

try <- levens("Culm_diam")
summary(try)
hsd <- TukeyHSD(try)

LABELS <- generate_label_df(hsd, 'df$Elevation')

output <- data.frame(LABELS)
g <- univariate_boxplot_trait_data("Culm_diam") +
    ylab("Culm diameter (mm)") +
    geom_text(data = output, 
              aes(x = treatment, y = rep(1.1*max(func_traits$Culm_diam), 3), label = Letters),
              size = 6)

#inflor

try <- levens("Inflor_length")
summary(try)
hsd <- TukeyHSD(try)

LABELS <- generate_label_df(hsd, 'df$Elevation')

output <- data.frame(LABELS)
h <- univariate_boxplot_trait_data("Inflor_length") +
    ylab("Inflorescence length (mm)") +
    geom_text(data = output, 
              aes(x = treatment, y = rep(1.1*max(func_traits$Inflor_length), 3), label = Letters),
              size = 6)
#spike

try <- levens("Spikelet_no")
summary(try)
hsd <- TukeyHSD(try)

LABELS <- generate_label_df(hsd, 'df$Elevation')

output <- data.frame(LABELS)
i <- univariate_boxplot_trait_data("Spikelet_no") +
    ylab("Spikelet number (mm)") +
    geom_text(data = output, 
              aes(x = treatment, y = rep(1.1*max(func_traits$Spikelet_no), 3), label = Letters),
              size = 6)
#Pedicel

try <- levens("Pedicel_length_epu")
summary(try)
hsd <- TukeyHSD(try)

LABELS <- generate_label_df(hsd, 'df$Elevation')

output <- data.frame(LABELS)
j <- univariate_boxplot_trait_data("Pedicel_length_epu") +
    ylab("Pedicel length (mm)") +
    geom_text(data = output, 
              aes(x = treatment, y = rep(1.1*max(func_traits$Pedicel_length_epu), 3), label = Letters),
              size = 6)

#Glume outer

try <- levens("Glume_length_outer")
summary(try)
hsd <- TukeyHSD(try)

LABELS <- generate_label_df(hsd, 'df$Elevation')

output <- data.frame(LABELS)
k <- univariate_boxplot_trait_data("Glume_length_outer") +
    ylab("Outer glume length (mm)") +
    geom_text(data = output, 
              aes(x = treatment, y = rep(1.1*max(func_traits$Glume_length_outer), 3), label = Letters),
              size = 6)

#glume inner

try <- levens("Glume_length_inner")
summary(try)
hsd <- TukeyHSD(try)

LABELS <- generate_label_df(hsd, 'df$Elevation')

output <- data.frame(LABELS)
l <- univariate_boxplot_trait_data("Glume_length_inner") +
    ylab("Inner glume length (mm)") +
    geom_text(data = output, 
              aes(x = treatment, y = rep(1.1*max(func_traits$Glume_length_inner), 3), 
                  label = Letters),
                  size = 6)

#spike width

try <- levens("Spikelet_width")
summary(try)
hsd <- TukeyHSD(try)

LABELS <- generate_label_df(hsd, 'df$Elevation')

output <- data.frame(LABELS)
m <- univariate_boxplot_trait_data("Spikelet_width") +
    ylab("Spikelet width (mm)") +
    geom_text(data = output, 
              aes(x = treatment, y = rep(1.1*max(func_traits$Spikelet_width), 3), 
                  label = Letters),
                  size = 6)
#glume inner

try <- levens("Spikelet_length")
summary(try)
hsd <- TukeyHSD(try)

LABELS <- generate_label_df(hsd, 'df$Elevation')

output <- data.frame(LABELS)
n <- univariate_boxplot_trait_data("Spikelet_length") +
    ylab("Spikelet length (mm)") +
    geom_text(data = output, 
              aes(x = treatment, y = rep(1.1*max(func_traits$Spikelet_length), 3), 
                  label = Letters),
                  size = 6)
#lemma outer

try <- levens("Lemma_outer_length")
summary(try)
hsd <- TukeyHSD(try)

LABELS <- generate_label_df(hsd, 'df$Elevation')

output <- data.frame(LABELS)
o <- univariate_boxplot_trait_data("Lemma_outer_length") +
    ylab("Outer lemma length (mm)") +
    geom_text(data = output, 
              aes(x = treatment, y = rep(1.1*max(func_traits$Lemma_outer_length), 3), 
              label = Letters),
              size = 6)
#lemma inner

try <- levens("Lemma_inner_length")
summary(try)
hsd <- TukeyHSD(try)

LABELS <- generate_label_df(hsd, 'df$Elevation')

output <- data.frame(LABELS)
p <- univariate_boxplot_trait_data("Lemma_inner_length") +
    ylab("Inner lemma length (mm)") +
    geom_text(data = output, 
              aes(x = treatment, y = rep(1.1*max(func_traits$Lemma_inner_length), 3), 
              label = Letters),
              size = 6)


ggpubr::ggarrange(a,b,c,d,e,f,g,h,i,j,k,l,m,n,o,p, align = "v", common.legend = T, legend = "right")
```



```{r codeml_pairs, fig.cap="Pairwise comparisons between taxa from codeml (Goldman & Yang 1994) (runmode -2, model = 0)"}

pairs <- read.csv(here("Data", "pairwise.csv"))

pairs <- pairs %>% mutate(Elev = 
                     case_when(seq %in% c(1:19, 31) & seq.1 %in% c(1:19, 31) ~ "Lowlands",
                               seq %in% c(21,22,28, 30, 34, 37,40) & seq.1 %in% c(21,22,28, 30, 34, 37,40) ~ "Ramosa",
                               seq %in% c(20,23:27, 29, 33, 35, 38, 39) & seq.1 %in% c(20,23:27, 29, 33, 35, 38, 39) ~ "Setacea"
                     )) %>% na.omit()

my_theme <- theme_bw() +
    theme(panel.grid.minor = element_blank(),
          panel.grid = element_blank(),
          legend.position = "none",
          legend.background = element_blank(),
          #axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          #axis.ticks.x = element_blank(),
          axis.title.y = element_text(size = 17),
          legend.text = element_text(size = 15)
    )

dn <- ggplot(pairs, aes(Elev, dN)) +
    geom_boxplot(aes(fill = Elev)) +
    my_theme +
    scale_fill_manual(values=wes_palette(n=3, name="Moonrise2"))


ds <- ggplot(pairs, aes(Elev, dS)) +
    geom_boxplot(aes(fill = Elev)) +
    my_theme +
    scale_fill_manual(values=wes_palette(n=3, name="Moonrise2"))

dnds <- ggplot(pairs, aes(Elev, dN.dS)) +
    geom_boxplot(aes(fill = Elev)) +
    my_theme +
    scale_fill_manual(values=wes_palette(n=3, name="Moonrise2"))

#free ratios


free_rates <- read.csv(here("Data", "free_rates.csv"))

free_rates <- free_rates %>% mutate(Elev = 
                     case_when(tip %in% c(1:19, 31) ~ "Lowlands",
                               tip %in% c(21,22,28, 30, 34, 37,40) ~ "Ramosa",
                               tip %in% c(20,23:27, 29, 33, 35, 38, 39) ~ "Setacea"
                     )) %>% na.omit()

dn_aov <- aov(free_rates$dN ~ free_rates$Elev)
ds_aov <- aov(free_rates$dS ~ free_rates$Elev)
dnds_aov <- aov(free_rates$dN_dS ~ free_rates$Elev)

dn_tuk <- TukeyHSD(dn_aov)
ds_tuk <- TukeyHSD(ds_aov)
dnds_tuk <- TukeyHSD(dnds_aov)


LABELS <- generate_label_df(dn_tuk, 'free_rates$Elev')
output <- data.frame(LABELS)

free_dn <- ggplot(free_rates, aes(Elev, dN)) +
    geom_boxplot(aes(fill = Elev)) +
    my_theme +
    scale_fill_manual(values=wes_palette(n=3, name="Moonrise2")) +
    geom_text(data = output, 
              aes(x = treatment, y = rep(1.1*max(free_rates$dN), 3), label = Letters),
              size = 6)

LABELS <- generate_label_df(ds_tuk, 'free_rates$Elev')
output <- data.frame(LABELS)
free_ds <- ggplot(free_rates, aes(Elev, dS)) +
    geom_boxplot(aes(fill = Elev)) +
    my_theme +
    scale_fill_manual(values=wes_palette(n=3, name="Moonrise2")) +
    geom_text(data = output, 
              aes(x = treatment, y = rep(1.1*max(free_rates$dS), 3), label = Letters),
              size = 6)


LABELS <- generate_label_df(dnds_tuk, 'free_rates$Elev')
output <- data.frame(LABELS)

free_dnds <- ggplot(free_rates, aes(Elev, dN_dS)) +
    geom_boxplot(aes(fill = Elev)) +
    my_theme +
  ylab(expression(omega))+
    scale_fill_manual(values=wes_palette(n=3, name="Moonrise2"))+
    geom_text(data = output, 
              aes(x = treatment, y = rep(1.1*max(free_rates$dN_dS), 3), label = Letters),
              size = 6)



ggpubr::ggarrange(dn, ds, dnds, free_dn, free_ds, free_dnds, align = "h", nrow = 2, ncol = 3)

```


```{r branch_mods}

branch_mods <- read.csv(here("Data", "branch_models.csv"))
kable(branch_mods,
      caption = "Parameter estimates, log-likelihood scores, and likelihood ratio test (LRT) P values obtained from branch models implemented in codeml, based on 44 concatenated exons. The null model fits a single $\\omega$ across the tree, while the remaining models estimate $\\omega$ for both foreground (the clade of interest) and background branches",
      format = "latex",
      align = "c",
      linesep = "",
      digits = 2,
      booktabs = T,
      escape = F,
      col.names = c("Model", "No. parameters", "Background $\\omega$",  "Foreground $\\omega$", "ln L", "$\\chi^2$ stat.", "LRT P value"))  %>%
  row_spec(12, hline_after = T) %>%  
  kable_styling(latex_options = c("striped", "hold_position"),
                    full_width = F,
                  position = "left") 
```
